---
title: 'Assignment 07: Web Scraping the FSB Directory'
author: "Lindsey Carothers"
date: "2026-10-05"
output: 
  html_document:
    code_download: true
    code_folding: show
    theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Currently, the Farmer School of Business has the following academic departments:


- Accountancy

- Economics

- Entrepreneurship

- Finance

- Information Systems & Analytics

- Marketing

- Management


Please write code that will produce and print a single data frame (or tibble) containing information on ALL departments and the following variables:

&emsp;[a] department name or abbreviation,

&emsp;[b] faculty/staff's name,
  
&emsp;[c] faculty/staff's position, and
  
&emsp;[d] faculty/staff's website.

Please print out the first few observations from the data frame using `head()`.


```{r}
departments = c(
  "Accountancy",
  "Economics",
  "Entrepreneurship",
  "Finance",
  "Information Systems & Analytics",
  "Marketing",
  "Management"
)

url_departments = c(
  "Accountancy",
  "Economics",
  "Entrepreneurship",
  "Finance",
  "Information%20Systems%20--%20Analytics",
  "Marketing",
  "Management"
)

fsb_df = data.frame()

for (i in 1:7) {
  
  url = paste0(
    "https://miamioh.edu/fsb/directory/?up=/query/all/all/",
    url_departments[i],
    "/all"
  )
  
  webpage = rvest::read_html(url)
  
  names = webpage |>
    rvest::html_elements("table strong a") |>
    rvest::html_text2()
  
  positions = webpage |>
    rvest::html_elements("table td i") |>
    rvest::html_text()
  
  websites = webpage |>
    rvest::html_elements("table strong a") |>
    rvest::html_attr("href")
  
  df = data.frame(
    department = departments[i],
    name = names,
    position = positions,
    website = websites
  )
  
  fsb_df = rbind(fsb_df, df)
}

head(fsb_df)
```


