---
title: "Class 07"
author: "Lindsey Carothers"
date: "2026-09-16"
output: 
html_document:
  toc: true
  toc_float: true
  number_sections: true
  code_folding: show
  code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Wrangling with dplyr
```{r, include=FALSE}
con = DBI::dbConnect(RSQLite::SQLite(), "../data/scout.db")
postings = DBI::dbReadTable(con, "scout_postings") |> tibble::as_tibble()
DBI::dbDisconnect(con)

jobs = postings |>
  dplyr::select(company, title, location_state, remote, category, posted_at, last_seen_at) |>
  dplyr::mutate(dplyr::across(dplyr::ends_with("_at"), as.Date),
                days_open = as.numeric(last_seen_at - posted_at))

jobs
```

## From SQL to dplyr

**Monday, Part 1: SQL, typed by you**
DBI::dbGetQuery(con, "
  SELECT company, COUNT(*) AS n
  FROM scout_postings
  WHERE remote = 1
  GROUP BY company
  ORDER BY n DESC
  LIMIT 10")
  
```{r}
# Today: dplyr, on the data frame from dbReadTable()
postings |>
  dplyr::filter(remote == 1) |>
  dplyr::count(company, sort = TRUE) |>
  head(10)
```

## The dplyr Verbs

dplyr::select(company, title, remote) #select

dplyr::filter(remote == 1, location_state == "OH") # where

dplyr::arrange(dplyr::desc(posted_at)) # order by

dplyr::mutate(days_open = last_seen - posted) # select ... as

dplyr::group_by(company) |> dplyr::summarize(n = dplyr::n()) # group by

dplyr::left_join(states, by = "location_state") # join


## A dplyr Pipeline, step by step
```{r}
jobs |>
  dplyr::select(company, location_state, remote, posted_at, last_seen_at) |>
  dplyr::filter(remote == 1) |>
  dplyr::mutate(days_open = as.numeric(as.Date(last_seen_at) - as.Date(posted_at))) |>
  dplyr::arrange(dplyr::desc(days_open)) |>
  dplyr::group_by(company) |>
  dplyr::summarize(n = dplyr::n(), avg_days = mean(days_open))
```

## select(): Choosing Columns
```{r}
con = DBI::dbConnect(RSQLite::SQLite(), "../data/scout.db")
postings = DBI::dbReadTable(con, "scout_postings") |> tibble::as_tibble()
DBI::dbDisconnect(con)

jobs = postings |>
  dplyr::select(company, title, location_state, remote, category, posted_at, last_seen_at)
jobs
```

## filter (): Choosing Rows
```{r}
c(one_test = jobs |> dplyr::filter(remote == 1) |> nrow(),
  both     = jobs |> dplyr::filter(remote == 1, location_state == "OH") |> nrow(),
  either   = jobs |> dplyr::filter(remote == 1 | category == "internship") |> nrow(),
  in_set   = jobs |> dplyr::filter(location_state %in% c("OH", "IN", "KY")) |> nrow())
```

## mutate() and arrange()
```{r}
jobs = jobs |>
  dplyr::mutate(days_open = as.numeric(as.Date(last_seen_at) - as.Date(posted_at)))

jobs |> dplyr::arrange(dplyr::desc(days_open)) |> dplyr::select(company, title, days_open)
```

## Selection Helpers and across()
```{r}
jobs = jobs |> dplyr::mutate(dplyr::across(dplyr::ends_with("_at"), as.Date))    # both dates at once
jobs |> dplyr::summarize(dplyr::across(dplyr::where(is.numeric), mean))
```

## group_by() and summarize()
```{r}
jobs |>
  dplyr::group_by(location_state) |>
  dplyr::summarize(postings = dplyr::n(), remote_share = mean(remote)) |>
  dplyr::arrange(dplyr::desc(postings))
```

# Demo

## Q1 -- which cat is remote friendly
To answer that, the logic would be as follows 
- Find the number of jobs per category
- Normalize the number of remote jobs by dividing by the total jobs
```{r}
jobs |> 
  # I want to move from unit of analysis being a job posting to category
  dplyr::group_by(category) |> 
  dplyr::summarise(
    postings = dplyr::n(),
    remote_share = mean(remote)
)
```
