---
title: "Class 10"
author: "Lindsey Carothers"
date: "2026-09-28"
output: 
  html_document:
    toc: true
    toc_float: true
    code_download: true
    code_folding: show
    theme: simplex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# When Can You Scrape

## `robots.txt`

In the chunk below, we used the [robotstxt package](https://www.google.com/goto?url=CAESZAHrOzAVD2a07KsmJF8fXZMF4AhNxipw7Sr0u7RPdCMMj1K5sw1Q1huBuWhfqUIx7jl8IoEMbvAZikGudHmVmjklRN_vu3pLfZlQjxmqkLjHw2RvC4AqDPKxNZb_YKsu5Wdm9uk) to check if we are allowed to write a program that scrapes the page.

```{r bulletin}
robotstxt::paths_allowed(
  paths = "/courses-instruction/isa/", 
  domain = "bulletin.miamioh.edu", 
  bot = "*"
  )
```
 
  > Returns TRUE. robots.txt is a request from the site owner, not a law. Respecting it is the first thing a careful scraper does.
  
### `robotstxt`: Wikipedia

```{r wiki}
robotstxt::paths_allowed(
  paths = "/wiki/Miami_RedHawks_football",
  domain = "en.wikipedia.org",
  bot = "*"
)
```

### `robotstxt`: LinkedIn

```{r linkedin}
robotstxt::paths_allowed(
  paths = "/jobs/",
  domain = "www.linkedin.com/",
  bot = "*"
)

# in some cases, the robotstxt provided inconsistent answers (not parsing the document correctly)
  # typically you get a warning with the answer 
  # solution: drop the trailing slash from the URL or check with the robots.txt (ground truth)

robotstxt::paths_allowed(
  paths = "/jobs/",
  domain = "www.linkedin.com",
  bot = "*"
)
```

### `robotstxt`: Indeed

```{r indeed}
robotstxt::paths_allowed(
  paths = "/jobs and /viewjob",
  domain = "www.indeed.com",
  bot = "*"
)
```


# How a Web Page Is Built

# Scraping with rvest

```{r}
url = "https://en.wikipedia.org/wiki/Miami_RedHawks_football"

# permission
robotstxt::paths_allowed(
  paths = "/wiki/Miami_RedHawks_football", 
  domain = "en.wikipedia.org", 
  bot = "*"
  )

# load
page = rvest::read_html(url)

# select
items = page |> rvest::html_elements("table.wikitable")

# extract
## extract table (a list of data frames)
items |> rvest::html_table()

## extract text (a character vector of text)
items |> rvest::html_text2()

## extract a character vector of attribute values
items |> rvest::html_attr("href")

```

# Scrape  Miami's Bulletin for ISA

```{r}
# checking whether we are allowed to scrape this info

robotstxt::paths_allowed(
  paths = "/courses-instruction/isa/", 
  domain = "bulletin.miamioh.edu", 
  bot = "*"
  )
```

```{r}
# I will load the backend of the HTML into R using 
  # [a] rvest::read_html() | sometimes rvest::read_html_live() if the page has some dynamic content
  # [b] in some cases, we will craft this as a request as shown in slides httr2::request() > .... > httr2::resp_body_html()

isa_courses_in_r = rvest::read_html("https://bulletin.miamioh.edu/courses-instruction/isa/")
```

```{r}
# we will extract the course title, which we can do in a single pipeline given that they look the same on the rendered web page in chrome 
  # (bolded, same font size and font type)

# CSS selector was obtained by:
  # highlighting what we wanted to pull (ISA 401 course title) in Chrome  
    # right click -> inspect -> right click -> copy -> copy selector

# find isa 401
isa_courses_in_r |> 
  rvest::html_elements("#sc_sccoursedescs > div:nth-child(24) > p.courseblocktitle > strong")

isa_courses_in_r |> 
  rvest::html_elements("#sc_sccoursedescs > div > p.courseblocktitle > strong") |> 
  rvest::html_text2() -> isa_titles

# course descriptions [using selector gadget, which does not always work]
isa_courses_in_r |> 
  rvest::html_elements(".courseblockdesc") |> 
  rvest::html_text2() -> isa_desc

df = data.frame(title = isa_titles, description = isa_desc)
```


















