---
title: "Class 07"
author: "Lindsey Carothers"
date: "2026-09-16"
output: 
html_document:
  toc: true
  toc_float: true
  number_sections: true
  code_folding: show
  code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Wrangling with dplyr
```{r, include=FALSE}
con = DBI::dbConnect(RSQLite::SQLite(), "../data/scout.db")
postings = DBI::dbReadTable(con, "scout_postings") |> tibble::as_tibble()
DBI::dbDisconnect(con)

jobs = postings |>
  dplyr::select(company, title, location_state, remote, category, posted_at, last_seen_at) |>
  dplyr::mutate(dplyr::across(dplyr::ends_with("_at"), as.Date),
                days_open = as.numeric(last_seen_at - posted_at))

jobs
```

## From SQL to dplyr

**Monday, Part 1: SQL, typed by you**
DBI::dbGetQuery(con, "
  SELECT company, COUNT(*) AS n
  FROM scout_postings
  WHERE remote = 1
  GROUP BY company
  ORDER BY n DESC
  LIMIT 10")
  
```{r}
# Today: dplyr, on the data frame from dbReadTable()
postings |>
  dplyr::filter(remote == 1) |>
  dplyr::count(company, sort = TRUE) |>
  head(10)
```

## The dplyr Verbs

dplyr::select(company, title, remote) #select

dplyr::filter(remote == 1, location_state == "OH") # where

dplyr::arrange(dplyr::desc(posted_at)) # order by

dplyr::mutate(days_open = last_seen - posted) # select ... as

dplyr::group_by(company) |> dplyr::summarize(n = dplyr::n()) # group by

dplyr::left_join(states, by = "location_state") # join


## A dplyr Pipeline, step by step
```{r}
jobs |>
  dplyr::select(company, location_state, remote, posted_at, last_seen_at) |>
  dplyr::filter(remote == 1) |>
  dplyr::mutate(days_open = as.numeric(as.Date(last_seen_at) - as.Date(posted_at))) |>
  dplyr::arrange(dplyr::desc(days_open)) |>
  dplyr::group_by(company) |>
  dplyr::summarize(n = dplyr::n(), avg_days = mean(days_open))
```

## select(): Choosing Columns
```{r}
con = DBI::dbConnect(RSQLite::SQLite(), "../data/scout.db")
postings = DBI::dbReadTable(con, "scout_postings") |> tibble::as_tibble()
DBI::dbDisconnect(con)

jobs = postings |>
  dplyr::select(company, title, location_state, remote, category, posted_at, last_seen_at)
jobs
```

## filter (): Choosing Rows
```{r}
c(one_test = jobs |> dplyr::filter(remote == 1) |> nrow(),
  both     = jobs |> dplyr::filter(remote == 1, location_state == "OH") |> nrow(),
  either   = jobs |> dplyr::filter(remote == 1 | category == "internship") |> nrow(),
  in_set   = jobs |> dplyr::filter(location_state %in% c("OH", "IN", "KY")) |> nrow())
```

## mutate() and arrange()
```{r}
jobs = jobs |>
  dplyr::mutate(days_open = as.numeric(as.Date(last_seen_at) - as.Date(posted_at)))

jobs |> dplyr::arrange(dplyr::desc(days_open)) |> dplyr::select(company, title, days_open)
```

## Selection Helpers and across()
```{r}
jobs = jobs |> dplyr::mutate(dplyr::across(dplyr::ends_with("_at"), as.Date))    # both dates at once
jobs |> dplyr::summarize(dplyr::across(dplyr::where(is.numeric), mean))
```

## group_by() and summarize()
```{r}
jobs |>
  dplyr::group_by(location_state) |>
  dplyr::summarize(postings = dplyr::n(), remote_share = mean(remote)) |>
  dplyr::arrange(dplyr::desc(postings))
```

# Demo

Which postings are remote, and does it depend on the job category, the state, or the kind of role? Typed live; you say the verb before I type it.
 
 
## Q1 -- Which category is remote friendly

One verb pair. The federal postings come from USAJobs and are on site by policy; internships almost never; so the remote question is a full-time question.

```{r}
# Find the number of jobs per category
# Normalize the number of remote jobs by dividing by the total jobs

jobs |> 
  # I want to move from unit of analysis being a job posting to category
  dplyr::group_by(category) |> 
  dplyr::summarise(
    postings = dplyr::n(),
    remote_share = mean(remote)
)
```

## Q2 -- Among full-time postings, which states have the highest remote share?

Keep one category, collapse to one row per state, keep the states with enough postings to trust a share, sort. Four verbs, one chain; each one answers a "but" from the room.

```{r}
jobs |> 
  dplyr::group_by(category, location_state) |> 
  dplyr::summarise(
    postings = dplyr::n(),
    remote_share = mean(remote) |> round(digits=2)
  ) |> 
  dplyr::arrange(dplyr::desc(remote_share)) |> 
  na.omit() # to drop the location_state = NA 

# Alternative way
jobs |> 
  dplyr:: filter(category == 'fulltime') |> 
  dplyr::group_by(location_state) |> 
  dplyr::summarise(
    postings = dplyr::n(),
    remote_share = mean(remote) |> round(digits=2)
  ) |> 
  dplyr::arrange(dplyr::desc(remote_share)) |> 
  na.omit() # to drop the location_state = NA 
```

## Q3 -- Does the kind of role matter?

The title column is free text, so the first step is a new column that reads it: dplyr::case_when() turns a set of tests into labels, in order, first match wins. Then the same pair of verbs as tab 1, plus how long each role stays posted.

```{r}

```

# Joins
```{r}
states = tibble::tibble(location_state = state.abb,
                        state_name = state.name,
                        region = as.character(state.region))

jobs |>
  dplyr::left_join(states, by = "location_state") |>
  dplyr::group_by(region) |>
  dplyr::summarize(postings = dplyr::n(), remote_share = mean(remote)) |>
  dplyr::arrange(dplyr::desc(postings))
```

## Anti-joins
```{r}
jobs |> dplyr::count(location_state) |> dplyr::anti_join(states, by = "location_state")
```

# Airbnb listings

```{r}
con = DBI::dbConnect(RSQLite::SQLite(), "../data/midwest_airbnb.db")
listings = DBI::dbReadTable(con, "listings") |>
  tibble::as_tibble()
DBI::dbDisconnect(con)
```
1. How many listings per city?
```{r}
listings |>  
  dplyr::group_by(city) |> 
  dplyr::summarise(count = dplyr::n())

# short cut using count
listings |> 
  dplyr::count(city)

# alternatively, you can count the number of times each city occurred in the city column
table(listings$city) # returns named vector
```

2. For entire homes (room_type == "Entire home/apt") that sleep four or more, what is the median price in each city?
```{r}
listings |> 
  dplyr::filter(room_type == "Entire home/apt" & accommodates >= 4) |> 
  dplyr::group_by(city) |> 
  dplyr::summarise(median_price = median(price))
```

3. Add price_per_person = price / accommodates. Which three listings are cheapest per person? Believe them?
```{r}
listings |> 
  dplyr::mutate(price_per_person = price / accommodates) |> 
  dplyr::select(city, host_name, name, price, accommodates, price_per_person) |> 
  dplyr::arrange(price_per_person) |> 
  head(3)
```

4. Which five host_names have the most listings?
```{r}
# host names are not unique, host_ids are

listings |> 
  dplyr::count(host_name, sort = T)

# this code finds unique hosts through host_id
listings |> 
  dplyr::count(host_name, host_id, sort = T)
```





