---
title: 'Class 11: Scraping Multiple Webpages'
author: "Lindsey Carothers"
date: "2026-09-30"
output: 
  html_document:
    toc: true
    toc_float: true
    code_download: true
    code_folding: show
    theme: simplex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Learning Objectives

- Be able to scrape multiple webpages (from the same site), using a 'for' loop type approach
- Be able to scrape multiple webpages (from the same site) using the [purrr package](https://purrr.tidyverse.org/), which is a functional programming approach
  - **Note**: In this package we will create our own function 
  
## Re-examining the bulletin example

Let's try to expand on the example from last class to include the following: 
[ACC](https://bulletin.miamioh.edu/courses-instruction/acc/), 
[BUS](https://bulletin.miamioh.edu/courses-instruction/bus/),
[ISA](https://bulletin.miamioh.edu/courses-instruction/isa/), and
[MKT](https://bulletin.miamioh.edu/courses-instruction/mkt/).

**COMMENT**: You can repeat your code a few time with changing the URL. this will work for a few links but would be quite tedious for large amount of scraping. We will follow **DRY Principle**, which stands for **Don't Repeat Yourself**.

```{r bulletin}
dept_abbrv = c('acc', 'bus', 'isa', 'mkt')

courses_df = data.frame() # initializing a container to store stuff in from the loop

# for (variable in vector) {}

for (abbrv in dept_abbrv) {
  cat(rep("-", 50), "\n")
  print(abbrv)
  
  bulletin_url = paste0(
    'https://bulletin.miamioh.edu/courses-instruction/', abbrv, '/'
    )

  # Intentionally making this an R object instead of a string
    # we can reuse the rest of our code with only changing the URL
  webpage = rvest::read_html(bulletin_url)

  webpage |> 
    rvest::html_elements("strong") |> # get all the info that's in 'strong' on the webpage
    rvest::html_text2() -> titles # cleans out the html text

  webpage |> 
    rvest::html_elements("#sc_sccoursedescs > div > p.courseblockdesc") |> 
    rvest::html_text2() -> descs

  df = data.frame(title = titles, desc = descs)
  courses_df = rbind(courses_df, df)
}
```

## 'purrr' approach for the same problem

This is a functional programming approach, and we will write our own function. In R, we use function called 'function' to create a function. as opposed to Python, R functions return a single object.

```{r}
# This function takes an abbrv as an input and returns a data_frame as an output

scrape_mu_bulletin = function(abbrv){
  bulletin_url = paste0(
    'https://bulletin.miamioh.edu/courses-instruction/', abbrv, '/'
    )

  webpage = rvest::read_html(bulletin_url)

  webpage |> 
    rvest::html_elements("p.courseblocktitle") |> 
    rvest::html_text2() -> titles 

  webpage |> 
    rvest::html_elements("#sc_sccoursedescs > div > p.courseblockdesc") |> 
    rvest::html_text2() -> descs

  df = data.frame(title = titles, desc = descs)
  return(df)
}

test1 = scrape_mu_bulletin('mgt')
test2 = scrape_mu_bulletin('fin')
test3 = scrape_mu_bulletin('eco')

# to use functions from purrr, we will use a map_*(), a rmap2_*(), or a pmap_*()
  # map uses a function that takes a single input (or you are only changing a single input)
  # map2 uses a function that takes two inputs
  # pmap >= 3
# after the _, you put the type of object you are getting back

# .x is the vector that we will iterate over
  # and .f is the function we use
courses_df_from_purrr = purrr::map_df(.x = dept_abbrv, .f = scrape_mu_bulletin)
```

## Scrape data from the website [planecrashinfo](https://www.planecrashinfo.com/)

Contains a database of plane crashes by year.

```{r, eval = FALSE}
year = 2025

crash_url = paste0(
  "https://www.planecrashinfo.com/", year, "/", year, ".htm")
  
crash_page = rvest::read_html(crash_url)

crash_tables <- crash_page |> 
  rvest::html_elements("table") |> 
  rvest::html_table() 
```







