---
title: "Class 06"
author: "Lindsey Carothers"
date: "2026-09-14"
output:
  html_document:
    toc: TRUE
    toc_float: true
    code_download: true
    code_folding: show
    theme: simplex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# HW 04 Review

```{r hw04}
con = DBI::dbConnect(RSQLite::SQLite(), "../data/midwest_airbnb.db")
listings = DBI::dbReadTable(con, "listings")
```

# Kahoot Review

- campus["Farmer"] gives you a shorter list, not a vector
  - a vector contains elements that are all the same type, a list contains different types pf elements
  - you have to use two square brackets in order to return a data frame from a data frame or vector from a vector
  
- we downloaded occupation.xlsx instead of reading it from its web address because read.excel() needs a file on disk

- the 'skip = 1' in a 'read.excel()...' skips the title row above the data

- fromJSON() quotes every value on the Cincinnati crash data, so R stores them as text

- fromJSON() is better to use than read_json()
  - read_csv() is better to use than read.csv()
  
```{r review_q7and8}
run = readr::read_csv(
  "https://isa401-find-your-seat.vercel.app/api/export?section=A&run=13")
dplyr::glimpse(run)
```

# Part 1

## A Database in a File
```{r part1}
# package::function()
con = DBI::dbConnect(RSQLite::SQLite(),
                     "../data/scout.db")
DBI::dbListTables(con)
postings = DBI::dbReadTable(con, "scout_postings")
DBI::dbDisconnect(con)
dplyr::glimpse(postings[, c("title", "company",
  "location_state", "remote", "skills_json")])
```

## JSON Inside a Cell
```{r}
first_cell = postings$skills_json[1]
cat(strtrim(first_cell, 60))   # base R: first 60 characters

skills = jsonlite::fromJSON(first_cell)
skills
```

## Ask the Database a Question in SQL
```{r}
con = DBI::dbConnect(RSQLite::SQLite(),
                     "../data/scout.db")
top_remote = DBI::dbGetQuery(con, "
  SELECT company, COUNT(*) AS n
  FROM scout_postings
  WHERE remote = 1
  GROUP BY company
  ORDER BY n DESC
  LIMIT 10")
DBI::dbDisconnect(con)
top_remote
```

# Part 2: Your First App










